TacticToe: Learning to Prove with Tactics

### 2018-03-20 tactictoe_jv.tex
integrating Josef's comments

\def\holfour{\textsf{HOL4}\xspace}
\def\isabelle{\textsf{Isabelle}\xspace}
\def\hollight{\textsf{HOL Light}\xspace}
\def\mizar{\textsf{Mizar}\xspace}
\def\coq{\textsf{Coq}\xspace}
\def\ocaml{\textsf{OCaml}\xspace}
\def\vampire{\textsf{Vampire}\xspace}
\def\eprover{\textsf{Eprover}\xspace}
\def\zthree{\textsf{z3}\xspace}

@@

\begin{abstract}
We implement a tactical prover TacticToe on top of the HOL4 interactive theorem
prover. TacticToe learns from human proofs which mathematical technique is
the most suitable in each proof situation.
This knowledge is then used by a Monte Carlo tree search algorithm to
explore the most promising tactic-level proof paths.
On a single CPU, with a time limit of 60 seconds, TacticToe proves 66.4\% of
7164 theorems in HOL4's standard library whereas
Eprover with auto-schedule solves 34.5\%. The success rate rises to 69.0\% by 
combining the results of TacticToe and Eprover.
\end{abstract}

@@

\paragraph{Lists of Goals as Output of Tactics}
A dataset of tactic outputs is compiled during orthogonalization (see 
Section~\ref{sec:ortho}).
Some elements of this set will be considered positive examples (see 
Section~\ref{sec:evaluation}).
Lists of goals that are a subset or equal to the result of a recorded tactic on 
a target goal are considered positive because if a set of goals is solvable 
then a subset of it is too. During proof search, given a list of
goals $l$ created by a tactic, we select a set of lists of goals that are most
similar to $l$. The ratio of positive examples in the selection gives us a
prior evaluation for MCTS.

@@

We estimate the similarity (or co-distance) between two objects $o_1$ and $o_2$
through their respective feature sets $f_1$ and $f_2$.
The estimation is based on the frequency of the features in the intersection of
$f_1$ and $f_2$. A good rule of thumb is that the rarer the shared features
are, the more similar two goals should be. The relative rarity of each feature
can be estimated by calculating its TF-IDF weight~\cite{Jones72astatistical}.
We additionally raise these weighs to the sixth power giving
even more importance
to rare features. This modification was found experimentally optimal in
\cite{ckju-pxtp13}.

The first co-distance $sim_1$ sums the weight of all shared
features to compute the total score.
In a second co-distance $sim_2$, we additionally take
the total number of features into account, to reduce the seemingly unfair
advantage of big
feature sets in the first scoring function.
\[sim_1 (o_1, o_2) = {\sum\nolimits_{\,f \in f_0 \cap
<f_1}{\text{tfidf}(f)^{6}}}\]
\[sim_2 (o_1, o_2) = \frac{{\sum\nolimits_{\,f \in f_0 \cap
f_1}{\text{tfidf}(f)^{6}}}}
{ln (e + \mathit{card} f_0 + \mathit{card} f_1)}\]

@@

\paragraph{Resolution}
The search ends when the algorithm reaches one of these 3 conditions:
\begin{itemize}
\item It finds a proof, i.e., the root node is solved. In this case,
the search returns a minimized and
prettified proof script (see Section~\ref{sec:proofdisplay}).
\item It saturates, i.e., there are no tactics to be applied to any open goal.
This occurs less than 10 times in the full-scale experiment.
This happens only at the beginning of \holfour library, where the training data
has very few tactics.
\item It exceeds a given time limit (default is set to 60 seconds).
\end{itemize}

@@

We now concentrate on the definition of a reasonable evaluation function for the
output of tactics. Theoretically, a list of goals is interesting if short proofs
are possible each the goals in the list, and the lengths of the proofs can be
estimated from previous proof searches. However, this
approach generates too much data and slows down the proof search. So, we only
collect outputs of tactics tested during orthogonalization. We
declare a list of goals $l$ to be positive if it is has been produced by the
winner of an orthogonalization competition.
The set of positive examples in
$\mathit{Predictions}^{\mathit{goal\_list}}_k(l)$ is denoted
$\mathit{Positive}^{\mathit{goal\_list}}_k(l)$.
We chose $k=10$ as a default evaluation radius in our experiments.
And we evaluate a node $a$ through the list of goals $G(a)$ it
contains using the prior evaluation function:
\begin{align*}
\mathit{PriorEvaluation}_k (a) &=
  \frac{card\ \mathit{Positive}^{\mathit{goal\_list}}_k(G(a)}{k}\\
\end{align*}

@@

\holfour already integrates a superposition-based prover \metis for this purpose. It is already
recognized by the tactic selection mechanism in \tactictoe and its
premises can be predicted dynamically when in an abstracted form. Nevertheless,
we think that the performance of \tactictoe can be boosted by giving the ATP
\metis a special status. This arrangement consists of always predicting
premises for \metis, giving it a slightly higher timeout and
trying it first on each open goal. These modifications only induce an
linear overhead on the number nodes as \metis does not create
new goals.

The integration of an external prover is also possible. For this purpose we experiment
with
\eprover~\cite{eprover}.
Namely, we developed a tactic which expects a list of premises as an
argument
and calls \eprover on the goal together with the premises translated to a first-order problem. If \eprover
succeeds on the problem, the premises used in the external proof are used to
reconstruct the proof inside \holfour with \metis. Giving a special status
to external
provers is essential as external ATP calls do not appear in human
proof scripts. For external ATPs, we use an even higher timeout (5 seconds) and
a larger number of predicted premises (128 for \eprover) and also try them
first on open
goals. Since calls to ATPs are
computationally expensive, they are run in parallel and asynchronously. This
avoids slowing down \tactictoe's search loop. The number of asynchronous calls
to external provers that can be executed at the same time is limited by the
number of cores available to the user.

@@

\begin{definition}(Dependencies of a tactic-goal pair)\\
Let $\mathfrak{F}$ be the set of recorded tactic-goal pairs.
The dependencies $D_\infty$ of a tactic-goal pair $(t_0,g_0)$ are
inductively defined by:

\begin{align*}
D_0 &=_{def} \lbrace (t_0,g_0) \rbrace \\
D_{n+1} &=_{def} D_n \cup \lbrace (t,g)\in \mathfrak{F}\  |\ \exists
(t',g') \in D_n.\ g \in t'(g') \rbrace  \\
D_\infty &=_{def} \bigcup_{i \in \mathbb{N}} D_i\\
\end{align*}
\end{definition}

\begin{definition}(Dependencies of a theorem)\\
Let $\mathfrak{T}$ be the set of recorded theorems.
The dependencies of a theorem $t$ are the set of top-level theorems in
$\mathfrak{T}$ appearing in the proof of $t$. These dependencies are
recorded by tracking the use of top-level theorems through the inference
kernel~\cite{tgck-cpp15}.
\end{definition}

@@

If one of the tactics is failing then $t_1$ and $t_2$ are not comparable through
this relation.
Given the default subsumption on goals, $t_1$ subsumes $t_2$ on $g$ if and only
if $t_1(g)$ is a subset (modulo $\alpha$-equivalence) of $t_2(g)$.
\end{definition}

@@

% All in all, 
A database built with the orthogonalization process hus contains
$(\mathbb{O}(t,g),g)$ instead of $(t,g)$.

@@

\paragraph{Abstraction of Tactic Arguments}
The first step is to abstract arguments of tactics by replacing them by a
placeholder, creating a tactic with an unspecified argument.
Suppose we have a recorded tactic $t$. Since $t$ is a \sml code tree, we can
try to abstract any of the
\sml subterms. Let $u$ be a \sml subterm of $t$, and $h$ a placeholder in $u$, we
can substitute $u$ by $h$ to create an abstracted tactic.
We denote this intermediate tactic $t[h/u]$. By repeating the
process, multiple
arguments may be abstracted in the same tactic. Ultimately, the more
abstractions are performed the more general a tactic is, as many tactics
become instances of the same abstracted tactic. As a consequence, it becomes
harder and harder to predict suitable arguments. In our experiments, we create
one abstraction $\hat{t}$ for each tactic $t$ by abstracting all subterms of
type theorem list in $t$.

\paragraph{Instantiation of an Abstracted Tactic}
An abstracted tactic is not immediately applicable to a goal, since it contains
unspecified arguments. To apply an abstracted tactic $\hat{t}$, we first need
to
instantiate the placeholders inside $\hat{t}$. Because it is difficult to
build a general predictor effective on each type of argument, we manually
design different argument predictors for each type. Those predictors are
given a goal $g$ as input and return the best predicted argument for this
goal.

@@

Although the evaluation of each set of parameters on its own is fair,
the selection of the best strategy in Section~\ref{sec:tuning} should also be
considered as a learning process. To ensure the global fairness, the final
experiments in Section~\ref{sec:full_exp} runs the best strategy on the full
dataset which is about 30 times larger.

@@

\paragraph{Reconstruction}
Tactic proofs produced by \tactictoe during this experiment are all verifiable
in \holfour. By the design of the proof search, reconstruction of \tactictoe
proof succeeds, unless one of the tactics modifies the state of \holfour in a
way that changes the behavior of a tactic used in the final proof. This has not
occurred a single time in our experiments.
More details on how a proof is extracted from the search tree is given in
Section~\ref{sec:proofdisplay}.
Comparatively, we achieve a reconstruction rate of 95 percent for \eprover,
by calling \metis for 2.0 seconds with the set of theorems used
in \eprover's proof as argument.

@@

\paragraph{Reinforcement learning}
All our feature vectors have been learned form human proofs. We now can now
also add tactic-goal pairs that appears in \tactictoe's final proof. To prevent
duplication of effort, orthogonalization of those
tactics is essential to have a beneficial effect.
Since recording and re-proving are intertwined during evaluation, the
additional data is available for the next proof search.
The hope is that the algorithm will improve faster by learning from its own
discovered proofs than from the human proof
scripts~\cite{DBLP:conf/cade/Urban07}. Side experiments show that
this one shot reinforcement learning method increases \tactictoe's success rate
by less than a percent.

@@

In order to appreciate the difficulty of the evaluation set from a human
perspective, the length distribution of human proofs in the
validation set is shown in Figure~\ref{fig:distrib}.  It is clear from the
graph that most of the human proofs are short. The proofs found by \tactictoe
follow a similar
distribution. % This means that 
If \tactictoe finds a proof, there is about a 50
percent chance that it will be shorter than what a human would come up with.

@@

In Figure~\ref{fig:percentage}, we regroup the human proofs by their length to
see how well \tactictoe and \eprover cope with increased levels of
difficulty. As expected, the longer the original proof is, the harder it is
for \tactictoe and \eprover to re-prove the theorem on their own.
The performance of \tactictoe is compelling with more than half of the theorems
that required a proof of length six being reproven. It is also consistently
better than \eprover for any proof length.

@@

Even when developing one theory, the context is continually changing:
modules are opened and local identifiers are defined. Therefore, it is unlikely
that code extracted from proof scripts is interpretable in any other part of
\holfour without any post-processing.
To solve this issue, we recursively replace each local identifier by its
definition until we are able to write any expression with global identifiers
only. Similarly, we prefix each global identifier by its module.
We call this process \emph{globalization}. The result is a standalone \sml code
executable in any \holfour theory.  In the absence of side effects, it is
interpreted in the same way across theories.
Therefore, we can guarantee that the behavior of recorded stateless tactics
does not change. And because an update on a stateful tactic
generally increases its strength, our prediction algorithm based on past
examples is still reasonably effective in this case.

@@

In the next phase, we globalize identifiers of the tactic proofs.
Infix operators such as \texttt{THEN} need to be processed in a special way so
that they keep their infixity status after globalization. For simplicity,
the
globalization of infix operators is omitted in Example~\ref{ex:r1}.
In this example, the three main cases that can happen during the globalization
are depicted. The first one is the globalization of identifiers declared in
modules. The global identifiers \texttt{STRIP\_TAC} and
\texttt{ASM\_REWRITE\_TAC} are prefixed by their module \texttt{Tactic}. In
this way, they will still be interpretable whether \texttt{Tactic} was open or
not. The local identifier \texttt{LIST\_INDUCT\_TAC} is replaced by its
definition which happens to contain two global identifiers.
The previous paragraph describes the globalization for all identifiers except
local theorems.  We do not replace a local
theorem by its definition because we want to avoid unfolding tactic proofs into
other tactic proofs.
If the theorem is stored in the \holfour database available across \holfour
developments, we can obtain the theorem value by calling \texttt{DB.fetch}.
Otherwise the globalization process fails and the local theorem identifier is
kept unchanged. A recorded tactic with an unchanged local theorem as argument
is only interpretable inside the current theory.

>>>>

\def\holfour{\textsf{HOL4}\xspace}
\def\isabelle{\textsf{Isabelle}\xspace}
\def\hollight{\textsf{HOL Light}\xspace}
\def\mizar{\textsf{Mizar}\xspace}
\def\coq{\textsf{Coq}\xspace}
\def\ocaml{\textsf{OCaml}\xspace}
\def\vampire{\textsf{Vampire}\xspace}
\def\eprover{\textsf{E~prover}\xspace}
\def\zthree{\textsf{z3}\xspace}

@@

\begin{abstract}
We implement a tactical prover TacticToe on top of the HOL4 interactive theorem
prover. TacticToe learns from human proofs which mathematical technique is
the most suitable in each proof situation.
This knowledge is then used by a Monte Carlo tree search algorithm to
explore the most promising tactic-level proof paths.
On a single CPU, with a time limit of 60 seconds, TacticToe proves 66.4\% of
7164 theorems in HOL4's standard library whereas
Eprover with auto-schedule solves 34.5\%. The success rate rises to 69.0\% by 
combining the results of TacticToe and E~prover.
\end{abstract}

@@

\paragraph{Lists of Goals as Output of Tactics}
A dataset of tactic outputs is compiled during orthogonalization (see 
Section~\ref{sec:ortho}).
Some elements of this set will be considered positive examples (see 
Section~\ref{sec:evaluation}).
During proof search, given a list of
goals $l$ created by a tactic, we select a set of lists of goals that are most
similar to $l$. The ratio of positive examples in the selection gives us a
prior evaluation for MCTS.

@@

We estimate the similarity (or co-distance) between two objects $o_1$ and $o_2$
through their respective feature sets $f_1$ and $f_2$.
The estimation is based on the frequency of the features in the intersection of
$f_1$ and $f_2$. A good rule of thumb is that the rarer the shared features
are, the more similar two goals should be. The relative rarity of each feature
can be estimated by calculating its TF-IDF weight~\cite{Jones72astatistical}.
We additionally raise these weights to the sixth power giving
even more importance
to rare features. This modification was found experimentally optimal in
\cite{ckju-pxtp13}. \todoi {Cezary: Josef says sixth power is not in this 
citation}.

The first co-distance $sim_1$ sums the weight of all shared
features to compute the total score.
In a second co-distance $sim_2$, we additionally take
the total number of features into account, to reduce the seemingly unfair
advantage of big
feature sets in the first scoring function.
\[sim_1 (o_1, o_2) = {\sum\nolimits_{\,f \in f_0 \cap 
f_1}{\text{tfidf}(f)^{6}}}\]
\[sim_2 (o_1, o_2) = \frac{{\sum\nolimits_{\,f \in f_0 \cap
f_1}{\text{tfidf}(f)^{6}}}}
{ln (e + \mathit{card} f_0 + \mathit{card} f_1)}\]

@@

\paragraph{Resolution}
The search ends when the algorithm reaches one of these 3 conditions:
\begin{itemize}
\item It finds a proof, i.e., the root node is solved. In this case,
the search returns a minimized and
prettified proof script (see Section~\ref{sec:proofdisplay}).
\item It saturates, i.e., there are no tactics to be applied to any open goal.
This occurs less than 10 times in the full-scale experiment.
This happens only at the beginning of the \holfour library, where the training 
data
has very few tactics.
\item It exceeds a given time limit (default is set to 60 seconds).
\end{itemize}

@@

We now concentrate on the definition of a reasonable evaluation function for the
output of tactics. Theoretically, a list of goals is interesting if we can find 
a short proof for it. We could estimate the length of a proof from previous 
proof searches. However, this
approach generates too much data and slows down the proof search. So, we only
collect outputs of tactics tested during orthogonalization (see 
Section~\ref{sec:ortho}). We
declare a list of goals $l$ to be \textit{positive} if it is has been produced 
by the winner of an orthogonalization competition.
The set of positive examples in
$\mathit{Predictions}^{\mathit{goal\_list}}_k(l)$ is denoted
$\mathit{Positive}^{\mathit{goal\_list}}_k(l)$.
We chose $k=10$ as a default evaluation radius in our experiments.
And we evaluate a node $a$ through the list of goals $G(a)$ it
contains using the prior evaluation function:
\begin{align*}
\mathit{PriorEvaluation}_k (a) &=
  \frac{card\ \mathit{Positive}^{\mathit{goal\_list}}_k(G(a)}{k}\\
\end{align*}

@@

\holfour already integrates a superposition-based prover \metis~\cite{metis} 
for this purpose. It is already
recognized by the tactic selection mechanism in \tactictoe and its
premises can be predicted dynamically when in an abstracted form. Nevertheless,
we think that the performance of \tactictoe can be boosted by giving the ATP
\metis a special status. This arrangement consists of always predicting
premises for \metis, giving it a slightly higher timeout and
trying it first on each open goal. These modifications only induce an
linear overhead on the number nodes as \metis does not create
new goals.

The integration of an external prover is also possible. For this purpose we experiment
with
\eprover~\cite{eprover}.
Namely, we developed a tactic which expects a list of premises as an
argument
and calls \eprover on the goal together with the premises translated to a first-order problem. If \eprover
succeeds on the problem, the premises used in the external proof are used to
reconstruct the proof inside \holfour with \metis. Giving a special status
to external
provers is essential as external ATP calls do not appear in human
proof scripts. For external ATPs, we use an even higher timeout (5 seconds) and
a larger number of predicted premises (128 for \eprover) and also try them
first on open
goals. Since calls to ATPs are
computationally expensive, they are run in parallel and asynchronously. To 
update the MCTS values, the 
result of each thread is back-propagated in the tree after completion. This
avoids slowing down \tactictoe's search loop. The number of asynchronous calls
to external provers that can be executed at the same time is limited by the
number of cores available to the user.

@@

\begin{definition}(Dependencies of a tactic-goal pair)\\
Let $\mathfrak{F}$ be the set of recorded tactic-goal pairs.
The dependencies $D^*$ of a tactic-goal pair $(t_0,g_0)$ are
inductively defined by:

\begin{align*}
D_0 &=_{def} \lbrace (t_0,g_0) \rbrace \\
D_{n+1} &=_{def} D_n \cup \lbrace (t,g)\in \mathfrak{F}\  |\ \exists
(t',g') \in D_n.\ g \in t'(g') \rbrace  \\
D^* &=_{def} \bigcup_{i \in \mathbb{N}} D_i\\
\end{align*}
\end{definition}

\begin{definition}(Dependencies of a theorem)\\
Let $\mathfrak{R}$ be the set of recorded theorems.
The dependencies of a theorem $t$ are the set of top-level theorems in
$\mathfrak{R}$ appearing in the proof of $t$. These dependencies are
recorded by tracking the use of top-level theorems through the inference
kernel~\cite{tgck-cpp15}.
\end{definition}

@@

If one of the tactics is failing then $t_1$ and $t_2$ are not comparable through
this relation.
%Given the default subsumption on goals, $t_1$ subsumes $t_2$ on $g$ if and only
%if $t_1(g)$ is a subset (modulo $\alpha$-equivalence) of $t_2(g)$.
\end{definition}

@@

% All in all, 
A database built with the orthogonalization process thus contains
$(\mathbb{O}(t,g),g)$ instead of $(t,g)$.

@@

\paragraph{Abstraction of Tactic Arguments}
The first step is to abstract arguments of tactics by replacing them by a
placeholder, creating a tactic with an unspecified argument.
Suppose we have a recorded tactic $t$. Since $t$ is a \sml code tree, we can
try to abstract any of the
\sml subterms. Let $u$ be a \sml subterm of $t$, and $h$ a variable (playing 
the role of a placeholder), we
can replace $u$ by $h$ to create an abstracted tactic.
We denote this intermediate tactic $t[h/u]$. By repeating the
process, multiple
arguments may be abstracted in the same tactic. Ultimately, the more
abstractions are performed the more general a tactic is, as many tactics
become instances of the same abstracted tactic. As a consequence, it becomes
harder and harder to predict suitable arguments. In our experiments, we create
one abstraction $\hat{t}$ for each tactic $t$ by abstracting all subterms of
type theorem list in $t$.

\paragraph{Instantiation of an Abstracted Tactic}
An abstracted tactic is not immediately applicable to a goal, since it contains
unspecified arguments. To apply an abstracted tactic $\hat{t}$, we first need
to
instantiate the placeholders $h$ inside $\hat{t}$. Because it is difficult to
build a general predictor effective on each type of argument, we manually
design different argument predictors for each type. Those predictors are
given a goal $g$ as input and return the best predicted argument for this
goal.

@@

Although the evaluation of each set of parameters on its own is fair,
the selection of the best strategy in Section~\ref{sec:tuning} should also be
considered as a learning process. To ensure the global fairness, the final
experiment in Section~\ref{sec:full_exp} runs the best strategy on the full
dataset which is about 30 times larger.

@@

\paragraph{Reconstruction}
Tactic proofs produced by \tactictoe during this experiment are all verifiable
in \holfour. By the design of the proof search, reconstruction of \tactictoe
proof succeeds, unless one of the tactics modifies the state of \holfour in a
way that changes the behavior of a tactic used in the final proof. This has not
occurred a single time in our experiments. And a tactical proof generated by 
\tactictoe during the final experiment takes on average 0.37 seconds to replay.
More details on how a proof is extracted from the search tree is given in
Section~\ref{sec:proofdisplay}.
Comparatively, we achieve a reconstruction rate of 95 percent for \eprover,
by calling \metis for 2.0 seconds with the set of theorems used
in \eprover's proof as argument.

@@

\paragraph{Reinforcement learning}
All our feature vectors have been learned form human proofs. We now can now
also add tactic-goal pairs that appears in the last proof of \tactictoe. To 
prevent
duplication of effort, orthogonalization of those
tactics is essential to have a beneficial effect.
Since recording and re-proving are intertwined during evaluation, the
additional data is available for the next proof search.
The hope is that the algorithm will improve faster by learning from its own
discovered proofs than from the human proof
scripts~\cite{DBLP:conf/cade/Urban07}. Side experiments show that
this one shot reinforcement learning method increases \tactictoe's success rate
by less than a percent.

@@

In order to appreciate the difficulty of the evaluation set from a human
perspective, the length distribution of human proofs in the
validation set is shown in Figure~\ref{fig:distrib}.  It is clear from the
graph that most of the human proofs are short. The proofs found by \tactictoe
follow a similar
distribution.
If \tactictoe finds a proof, there is about a 50
percent chance that it will be shorter than what a human would come up with.

@@

In Figure~\ref{fig:percentage}, we regroup problems by the length of their 
human proof to measure how well \tactictoe and \eprover cope with increasing 
levels 
of difficulty. As expected, the longer the original proof is, the harder it is
for \tactictoe and \eprover to re-prove the theorem on their own.
The performance of \tactictoe is compelling with more than half of the theorems
that required a proof of length six being reproven. It is also consistently
better than \eprover for any proof length.

@@

Even when developing one theory, the context is continually changing:
modules are opened and local identifiers are defined. Therefore, it is unlikely
that code extracted from proof scripts is interpretable in any other part of
\holfour without any post-processing.
To solve this issue, we recursively replace each local identifier by its
definition until we are able to write any expression with global identifiers
only. Similarly, we prefix each global identifier by its module.
We call this process \emph{globalization}. The result is a standalone \sml code
executable in any \holfour theory.  In the absence of side effects, it is
interpreted in the same way across theories.
Therefore, we can guarantee that the behavior of recorded stateless tactics
does not change. Even with some stateful tactics, the prediction algorithm is  
effective because most updates on the state of \holfour increase the 
strength of stateful tactics.

@@

In the next phase, we globalize identifiers of the tactic proofs.
Infix operators such as \texttt{THEN} need to be processed in a special way so
that they keep their infixity status after globalization. For simplicity,
the
globalization of infix operators is omitted in Example~\ref{ex:r1}.
In this example, the three main cases that can happen during the globalization
are depicted. The first one is the globalization of identifiers declared in
modules. The global identifiers \texttt{STRIP\_TAC} and
\texttt{ASM\_REWRITE\_TAC} are prefixed by their module \texttt{Tactic}. In
this way, they will still be interpretable whether \texttt{Tactic} was open or
not. The local identifier \texttt{LIST\_INDUCT\_TAC} is replaced by its
definition which happens to contain two global identifiers.
The previous paragraph describes the globalization for all identifiers except
local theorems.  We do not replace a local
theorem by its tactic proof (which its \sml definition) because we want to 
avoid unfolding proofs inside other proofs.
If the theorem is stored in the \holfour database available across \holfour
developments, we can obtain the theorem value by calling \texttt{DB.fetch}.
Otherwise the globalization process fails and the local theorem identifier is
kept unchanged. A recorded tactic with an unchanged local theorem as argument
is only interpretable inside the current theory.
